The Emergent Simplicity of Galaxy Size

### 2017-09-29 paper/galsize_paper.tex
Tightened up abstract

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius of galaxies, $\rhalf.$ Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. We find that smaller galaxies cluster stronger relative to larger galaxies of the same stellar mass, a new result. We show that the $\rhalf-$dependence of galaxy clustering is largely driven by centrals being larger than satellite galaxies of the same halo mass. Stellar mass stripping of satellites has only a mild impact and is unlikely to be the origin of the characteristic magnitude and shape of the clustering signal. Much more successful models are based on the assumption that the mean galaxy profile {\em and its scatter} are correlated with the formation history of the dark matter halo profile. Based on our present findings, we predict that at fixed $\mhalo,$ galaxies with relatively shallow surface brightness profiles reside in relative low-concentration halos, and conversely. In the class of models that give the most faithful recovery of the observed clustering signal, galaxy $\rhalf$ is set by the size of the  virial radius at the time the subhalo reaches its peak halo mass, with log-normal scatter in $\rhalf$ that is strongly correlated with the halo scale radius $R_{\rm s}$ at that time. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}
%Using forward-modeling techniques based on {\tt Halotools}, we confirm previous results in Kravtsov (2013) that $\rhalf$ is well-described by a linear scaling relation with halo virial radius.

>>>>

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius of galaxies, $\rhalf.$ Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. Smaller galaxies cluster stronger relative to larger galaxies of the same stellar mass, a new result. We show that this clustering signal is largely driven by centrals being larger than satellite galaxies of the same halo mass. We build a simple model in which $\rhalf$ is set by the halo virial radius $\rvir$ at the time the halo reaches its peak mass, with scatter in $\rhalf$ that is strongly correlated with the halo scale radius $R_{\rm s}$ at that time. The $\rhalf-$dependent clustering of this model is strikingly similar to our SDSS measurements in magnitude, scale-dependence, and $\mstar-$dependence. Together with the result that stellar mass stripping of satellites has only a mild impact on $\rhalf-$dependent clustering, this suggests that the relative size of centrals and satellites is already in place at the time of satellite infall, and supports the notion that $L_{\ast}$ galaxy and halo profiles co-evolve across many Gyr of cosmic time. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}
